Exact Phylodynamics via Structured Markov Genealogy Processes

### 2023-07-06 ms.Rnw
edits to last proposition

The results so far allow us to compute expectations over random jumps and jump times.
We will have occasion to compute marginal expectations in the situation where some jump times are known.
The following result handles this case.
\AAK{[Not sure if this is stated correctly.]}

\begin{prop}
  Suppose $X_t$ is a continuous-time Markov process with state space $\Xspace$ and KFE as in \cref{eq:kfe}.
  Let $H_t$ be its history process and $E_t=\event(H_t)$, the set of its jump times to time $t$.
  Suppose $0<t_1<\dots<t_N\le{t}$ are fixed times and set $O=\{t_1,\dots,t_N\}$.
  Suppose $F$ is an $\Rp$-valued function of $H_t$ and $O$ such that
  \begin{equation*}
    F(H_t,O)=\prod_{e\in{E_t}}{Q\left(e,\predec{H}(e),H(e)\right)\,B\left(e,\predec{H}(e),H(e)\right)}
    \times\prod_{e\in{O}}{R\left(e,\predec{H}(e),H(e)\right)\,C\left(e,\predec{H}(e),H(e)\right)},
  \end{equation*}
  for some given measurable functions $B,C>{0}$, $Q,R\in\{0,1\}$.
  Suppose $w$ satisfies the filter equation
  \begin{equation*}
    \begin{aligned}
      \frac{\partial{w}}{\partial{t}}(t,x)
      =&\int{w(t,x')\,\beta(t,x',x)\,Q(t,x',x)\,B(t,x',x)\,\dd{x'}}
      -\int{w(t,x)\,\beta(t,x,x')\,Q(t,x,x')\,\dd{x'}}\\
      &-\sum_{e\in{O}}{\int{w(t,x')\,\delta(t,e)\,\,R(e,x',x)\,C(e,x',x)\,\dd{x'}}}
      -\sum_{e\in{O}}{\int{w(t,x)\,\delta(t,e)\,\,R(e,x,x')\,\dd{x'}}}\\
      &\qquad-\int{w(t,x)\,\beta(t,x,x')\,\left(1-Q(t,x,x')\right)\,\dd{x'}}.
    \end{aligned}
  \end{equation*}
  Then $\Expect{F(H_t,O)}=\int{w(t,x)\dd{x}}$.
\end{prop}
\begin{proof}
  Apply \cref{lemma:filtereq} with the driver generated by $\beta(t,x,x')=\alpha(t,x,x')+\sum_{e\in{O}}{\delta(t,e)\,\pi(x,x')}$.
\end{proof}

>>>>

The results so far allow us to compute expectations over random jumps and jump times.
We will have occasion to compute marginal expectations in the situation where some jump times are known.
The following result handles this case.

\begin{prop}
  Suppose $\X_t$ is a continuous-time Markov process with state space $\Xspace$ and KFE as in \cref{eq:kfe}.
  Let $\Hist_t$ be its history process.
  Suppose $0<t_1<\dots<t_N\le{t}$ are fixed times and set $E=\{t_1,\dots,t_N\}$.
  Let $B>{0}$ and $Q\in\{0,1\}$ be given measurable functions.
  Suppose $F_E$ is an $\Rp$-valued random variable such that
  \begin{equation*}
    \Expect{F_E\;\vert\;\Hist_t}={\prod_{e\in{E}}{B(e,\Xt_e,\X_e)\,Q(e,\Xt_e,\X_e)}}.
  \end{equation*}
  Let $\pi(x,x')$ be a given probability kernel and $Y_t$ be the Markov process generated by $\beta(t,x,x')=\left(\alpha(t,x,x')+\sum_{e\in{E}}{\delta(t,e)\,\pi(x,x')}\right)\,Q(t,x,x')$.
  Suppose $w$ satisfies the filter equation with driver $Y_t$, boost $B$, and decay $\lambda=\int{\alpha(t,x,x')\,\left(1-Q(t,x,x')\right)\,\dd{x'}}$.
  Then $\Expect{F_E}=\int{w(t,x)\,\dd{x}}$.
\end{prop}
\begin{proof}
  Apply \cref{lemma:filtereq}.
\end{proof}
